Deriving a precise Newtonian equation for the time it takes two separated bodies in an isolated system to collide due to the gravitational force

### 2022-10-30 paper/draft.tex
draft fixes

@@

Substituting this result in (TODO) and taking its square root results in

@@

@@

Similar to the reasoning for acceleration, we know that the velocity causes $r$ to become smaller, therefore we should be looking at the negative solution given by (TODO). Additionally rewriting it slightly results in

@@

@@

@@

@@

@@

Combining (todo: previous two equations) we get that the right side integral of (todO) can be rewritten as

@@

@@

And since we defined $w = tan(z)$, $z = tan^{-1}(w)$ which means (TODO) becomes

@@

Now, $tan^{-1}$ gives us an angle if we supply it with the ratio of opposite over adjacent, and if we supply this angle to the $sin$ function, we would get back the ratio of opposite over the hypotenuse, that is what $sin$ does. Can't we calculate this directly? We can, because we already supply $tan^{-1}$ with both the opposite and adjacent side with which we can calculate the hypotenuse. The adjacent side is just simply $1$, so therefore sine and by the same reasoning cosine, are given as follows

@@

We have the solution to our integral, so we can now just solve the definite integral for $\Delta T$ to get our final equation:

@@

Now, for the simplified case where we disregard the radii of the objects, $R_f$ would be $0$ upon collision, so that our equation in this case becomes

@@

@@

We will soon come back to this beautiful result when we are verifying our solution in the next section.

@@

which is exactly the same as the simplified result that we already found! That should give us enough confidence, but let us create a simple simulation program in Python so that we can compare the result of a simulation to the more complex solution that includes the radii of the objects.\\

@@

Below is the script that simulates the problem. It does this by calculating the acceleration, applying the acceleration to the velocity and applying the velocity to the distance. It calculates this every $0.001$ second (or really whatever you set the variable $dt$ to, until the objects collide. We use the Decimal class for enhanced precision. After the objects collided in the simulation, it prints the simulated time it took for the objects to collide and also the times it takes according to our equation. Then finally it prints the time it takes using Kepler's equation.

>>>>

    \label{eq:v_squared}

@@

Substituting this result in (\ref{eq:v_squared}) and taking its square root results in

@@

    \label{eq:v_pm_solution}

@@

Similar to the reasoning for acceleration, we know that the velocity causes $r$ to become smaller, therefore we should be looking at the negative solution given by (\ref{eq:v_pm_solution}). Additionally rewriting it slightly results in

@@

    \label{eq:primary_solution}

@@

    \label{eq:solution_w}

@@

    \label{eq:w_squared_add_one}

@@

    \label{eq:derivative_w_dz}

@@

Combining (\ref{eq:w_squared_add_one}) and (\ref{eq:derivative_w_dz}) we get that the right side integral of (\ref{eq:solution_w}) can be rewritten as

@@

    \label{eq:cos_squared}

@@

And since we defined $w = tan(z)$, $z = tan^{-1}(w)$ which means (\ref{eq:cos_squared}) becomes

@@

Now, $tan^{-1}$ gives us an angle if we supply it with the ratio of opposite over adjacent, and if we supply this angle to the $sin$ function, we would get back the ratio of opposite over the hypotenuse, that is what $sin$ does. Can we calculate the result of $sin$ directly? Yes, we can, because we already supply $tan^{-1}$ with both the opposite and adjacent side with which we can calculate the hypotenuse. The adjacent side is just simply $1$, so therefore sine and by the same reasoning cosine, are given as follows

@@

We have the solution to our integral, so we can now just substitute it in (\ref{eq:primary_solution}) and solve the definite integral for $\Delta T$ to get our final equation:

@@

So there we have it, our final result. Now, for the simplified case where we disregard the radii of the objects, $R_f$ would be $0$ upon collision, so that our equation in this case becomes

@@

    \label{eq:simplified_kepler}

@@

We will soon come back to this beautiful simplification when we are verifying our solution in the next section.

@@

which is exactly the same as the simplified result (\ref{eq:simplified_kepler}) that we already found! That should give us enough confidence, but let us create a simple simulation program in Python so that we can compare the result of a simulation to the more complex solution that includes the radii of the objects.\\

@@

Below is the script that simulates a hypothetical situation where the earth stops following its orbit and starts falling straight into the sun. The script does this by calculating the acceleration at the current distance, applying the acceleration to the velocity and applying the velocity to the distance. It calculates this every $0.001$ second (or really whatever you set the variable $dt$ to, until the objects collide. We use the Decimal class for enhanced precision. After the objects collide, it prints the time it took in our simulation and the time it should take according to our equation. Then finally it prints the time according to the Kepler's equation.
